Supplementary Material for ``Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)"

### 2024-01-27 ba_resubmit_template_files/supplement4.tex
bk eddits

	Due to the independence in the fast beta prior in \eqref{eqn:fast_beta_prior}, we can obtain the full conditional for $\bm{Z}$ through the full conditionals for each individual $Z_j$. Let $\Gamma_{.j}$ denote the random matrix of $n_A$ comparison vectors relating to an arbitrary record $B_j$, and let $\gamma_{.j}$ be a realization of $\Gamma_{.j}$. We have
	\begin{align}
		p(\bm{Z} | \gamma, \bm{m}, \bm{u}, \pi) &= \prod_{j=1}^{n_B} p(Z_j | \gamma_{.j}, \bm{m}, \bm{u}, \pi).
		%&\propto p(\Gamma| \bm{Z}, \bm{m}, \bm{u}) p(\bm{Z} | \pi) \\
		%&=\prod_{j=1}^{n_B} p(\Gamma_{.j}| Z_j, \bm{m}, \bm{u}) p(Z_j | \pi) \\
	\end{align}
{
\color{blue}{
Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
\begin{align}
	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
\end{align}
When $Z_j = q$ for some $q\leq n_A$, we have
\begin{align}
	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi,  Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
\end{align}
We multiply and divide by the $u$ parameters for the matching record pair to obtain
\begin{align}
	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
	&= w_{qj}  c_j.
\end{align}
We can divide the result of each case by $c_j$ to get
\begin{align}
	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
		w_{qj}, & q \leq n_A; \\
		1, &  q  = n_A + j.
	\end{cases}
\end{align}
Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
\begin{align}
	\label{eqn:z_full_conditional2}
	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
	\begin{cases} 
		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
		1 - \pi^{(s+1)}, & q  = n_A + j.
	\end{cases}
\end{align}
}
}

{
\color{red}{Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
	\begin{align}
		\mathcal{L}(Z_j = n_A + j| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
	\end{align}
	When $Z_j = q$ for some $q\leq n_A$, we have
	\begin{align}
		\mathcal{L}(Z_j = q| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
	\end{align}
	We multiply and divide by the $u$ parameters for the matching record pair to obtain
	\begin{align}
		\mathcal{L}(Z_j = q| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
		&= w_{qj}  c_j.
	\end{align}
	We can divide the result of each case by $c_j$ to get
	\begin{align}
		\mathcal{L}(Z_j | \bm{m}, \bm{u}, \pi, \Gamma_{.j}) \propto \begin{cases} 
			w_{qj}, & q \leq n_A; \\
			1, &  q  = n_A + j.
		\end{cases}
	\end{align}
	Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
	\begin{align}
		\label{eqn:z_full_conditional2}
		p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
		\begin{cases} 
			\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
			1 - \pi^{(s+1)}, & q  = n_A + j.
		\end{cases}
	\end{align}
}
}

>>>>

	Due to the independence in the fast beta prior in \eqref{eqn:fast_beta_prior}, we can obtain the full conditional for $\bm{Z}$ through the full conditionals for each individual $Z_j$. Let $\Gamma_{.j}$ denote the random matrix of $n_A$ comparison vectors relating to an arbitrary record $B_j$, and let $\gamma_{.j}$ be a realization of $\Gamma_{.j}$. We have
	\begin{align}
		p(\bm{Z} | \gamma, \bm{m}, \bm{u}, \pi) &= \prod_{j=1}^{n_B} p(Z_j | \gamma_{.j}, \bm{m}, \bm{u}, \pi).
		%&\propto p(\Gamma| \bm{Z}, \bm{m}, \bm{u}) p(\bm{Z} | \pi) \\
		%&=\prod_{j=1}^{n_B} p(\Gamma_{.j}| Z_j, \bm{m}, \bm{u}) p(Z_j | \pi) \\
	\end{align}
%{
%\color{blue}{
%Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
%\begin{align}
%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = n_A + j) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
%\end{align}
%When $Z_j = q$ for some $q\leq n_A$, we have
%\begin{align}
%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi,  Z_j = q) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
%\end{align}
%We multiply and divide by the $u$ parameters for the matching record pair to obtain
%\begin{align}
%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j = q) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
%	&= w_{qj}  c_j.
%\end{align}
%We can divide the result of each case by $c_j$ to get
%\begin{align}
%	p(\Gamma_{.j}| \bm{m}, \bm{u}, \pi, Z_j) \propto \begin{cases} 
%		w_{qj}, & q \leq n_A; \\
%		1, &  q  = n_A + j.
%	\end{cases}
%\end{align}
%Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) to obtain the full conditional
%\begin{align}
%	\label{eqn:z_full_conditional2}
%	p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
%	\begin{cases} 
%		\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
%		1 - \pi^{(s+1)}, & q  = n_A + j.
%	\end{cases}
%\end{align}
%}
%}

Following the observation of \cite{wortman2019}, when $B_j$ does not link to any record in $A$, the contribution to the likelihood is simply a product of $u$ parameters, which we will call $c_j$:
	\begin{align}
		\mathcal{L}(Z_j = n_A + j| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) = \prod_{i=1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} = c_j.
	\end{align}
	When $Z_j = q$ for some $q\leq n_A$, we have
	\begin{align}
		\mathcal{L}(Z_j = q| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) =\prod_{f=1}^{F}\prod_{l=1}^{L_f} m_{fl}^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i \neq q}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)}.
	\end{align}
	We multiply and divide by the $u$ parameters for the matching record pair to obtain
	\begin{align}
		\mathcal{L}(Z_j = q| \bm{m}, \bm{u}, \pi, \Gamma_{.j}) &= \prod_{f=1}^{F}\prod_{l=1}^{L_f} \left(\frac{m_{fl}}{u_{fl}}\right)^{I(\gamma_{qj}^f = l)I_{obs}(\gamma_{qj}^f)}  \prod_{i = 1}^{n_A}\prod_{f=1}^{F}\prod_{l=1}^{L_f} u_{fl}^{I(\gamma_{ij}^f = l)I_{obs}(\gamma_{ij}^f)} \\
		&= w_{qj}  c_j.
	\end{align}
%	We can divide the result of each case by $c_j$ to get
%	\begin{align}
%		\mathcal{L}(Z_j | \bm{m}, \bm{u}, \pi, \Gamma_{.j}) \propto \begin{cases} 
%			w_{qj}, & q \leq n_A; \\
%			1, &  q  = n_A + j.
%		\end{cases}
%	\end{align}
	Lastly, we multiply the likelihood by the fast beta prior in (\ref{eqn:fast_beta_prior}) and divide by $c_j$ to obtain the full conditional
	\begin{align}
		\label{eqn:z_full_conditional2}
		p\left(Z_j^{(s+1)}  = q|\gamma, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
		\begin{cases} 
			\frac{\pi^{(s+1)}}{n_A} w_{qj}^{(s+1)},  & q \leq n_A; \\
			1 - \pi^{(s+1)}, & q  = n_A + j.
		\end{cases}
	\end{align}
